The value of intermediation -- results so far

### 2022-07-22 src/paper/paper.tex
Add first version of the application section

>>>>

\section{Application}

Consider a market with a continuum of firms $a(t)$ ($t \in [0, 1]$), continuum of identical consumers and one platform $P$. The former two can only interact through the platform. Assume a measure $N \in (0, 1)$ of the firms are present on the platform. Then, consumers have the following quasilinear/CES utility function over money ($m$) and those firms' products:
\begin{align*}
    u(m, x(\cdot)) = m + \alpha \left[ \int_0^N x(t)^\rho \dt \right] ^ \frac{\gamma}{\rho}.
\end{align*}

\subsection{Timing of the game}

The game proceeds as follows. In the first stage, firms decide whether to create a product or not. Doing this entails a fixed cost of $C$. Those firms that choose not to invest obtain a payoff of 0 and are excluded from the remainder of the game. Next, the firms that made an investment can decide to join the platform at the cost of some entry fee $F^f$. (Note that the entry fee can be negative, in which case it is a subsidy from the platform to the firm.) After that, consumers decide to join or not at an entry fee of $F^c$. Finally, firms post their product prices, and produce and sell to consumers at marginal cost $c$.

The model heavily depends on how the entry fees are determined. The most widespread assumption in the literature is that they are unilaterally decided by the platform, and are essentially take it or leave it offers for the other players. The issue with this approach is that it generally leads to extreme conclusions. \textcolor{red}{(TODO: elaborate on this)} As opposed to this, in this paper I assume that firm entry fees are a result of some bargaining process between the firms and the platform, leading to an allocation in which the final profits of each player correspond to their Shapley values. User entry fees, on the other hand, are unilaterally decided by the platforms.

The value (characteristic function) of a coalition of the platform and any set of firms is defined as follows. Assume that the firms in the coalition enter the platform, while the others do not. Then, taking that decision as fixed, platforms set a profit-maximizing entry price for the consumers and firms also set own-profit-maximizing prices. That is, there is no collusion between the coalition members in terms of pricing. All they contract on are entry fees and decisions.

\subsection{Solution of the game}

Let us start from the last stage, and assume that firms and consumers have already made their entry decisions: a measure $N$ of firms and measure $M$ of consumers are on the platform. Furthermore, firms have already set their prices $p(t)$, $t \in [0, N]$. Then consumers face the following utility maximization problem:
\begin{align*}
    &u(m, x(\cdot)) = m + \alpha \left[ \int_0^N x(t)^\rho \dt \right] ^ \frac{\gamma}{\rho} \\
    &\text{s.t. } m + \int_0^N p(t) x(t) \dt = I.
\end{align*}
By following the usual solution method for CES utility functions, we obtain the following demand function:
\begin{align*}
    x(t) = M K p(t) ^ \frac{1}{\rho-1}
\end{align*}
where
\begin{align*}
    K = (\alpha \gamma) ^ {\frac{1}{1 - \rho} + \gamma - \rho} \left[ \int_0^N p(t) ^ \frac{\rho}{\rho - 1} \dt \right] ^ \frac{\rho - \gamma}{\rho (\gamma - 1)}.
\end{align*}
It implies that, as with the usual CES utility function, the optimal pricing strategy is independent on others' prices or the number of consumers, and is simply a constant markup over the marginal cost:
\begin{align*}
    p = \frac{c}{\rho}.
\end{align*}
The total profit each firm makes is then $\pi(N)$, while the total profit of all firms is $\Pi(n) = N\pi(N)$.

Consumers utility is from buying the products is then $u(N)$, which is increasing in $N$. As all consumers are identical, the platform optimally sets a user entry price of $F^u = u(N)$, so that it extracts all surplus from the consumers. Therefore, the total profits that a coalition of firms and the platform can obtain (assuming they individually maximize profits after the entry decisions are made) is
\begin{align*}
    v(N) = \Pi(N) + U(N).
\end{align*}
$v(N)$ is increasing in $N$, and its shape depends on the parameters of the utility function of the consumers and the marginal cost of the producers.

\subsection{Deciding firm entry fees}

Let us look at the second stage of the game: firms have made their investment decisions, and are now negotiating with the platform over the entry price. Let us assume that $T \in [0, 1]$ measure of firms have made an investment. $C$ is a fixed cost, therefore does not influence the rest of the game. Remember that entry fees will be set such that the final profits correspond to Shapley values. Let us denote these values by $\varphi_f(v, T)$ and $\varphi_P(v, T)$ for firms and the platform, respectively. As shown in the main results,
\begin{align*}
    \varphi_P(v, T) &= \int_0^1 v(tT) \dt \\
    \varphi_f(v, T) &= v(T) - \int_0^1 v(tT) \dt.
\end{align*}
In order to achieve these values, the firm entry fee will be set to $F^f = \varphi_P(v, T) - U(T)$.

Finally, looking at the first stage, firms will make the initial investment if and only if $\varphi_f(v, T) / T \geq C$. Therefore, the equilibrium measure of entrants is given by
\begin{align*}
    v(T) - \int_0^1 v(tT) = CT.
\end{align*}
